---
title: "GS Walk Through: Temperature"
author: "Grace Shiffrin"
date: "2024-04-08"
output: 
  html_document:
    toc: true
    highlight: zenburn
---

# Uploading packages and data
```{r package}
library(tidyverse)
library(dplyr)
library(ggplot2)
require("terra")
#library(lubridate)
#library(bipartite)
#library(lme4)
#library(GGally)
#library(rmarkdown)
#library(knitr)
```

Next, **setting my working directory** - this allows me to save to a hard drive that is connected to my device. If you only want to run through the code and don't want to save this specific file, you can remove this code chunk or place a # in from of the "setwd("D:/FS2_othr")". If you prefer one drive for saving, you will have to change this line to connect to your one drive.
```{r set working directory}
setwd("D:/gs4github/Temp")
```


**Importing the data (csv file).**  If you prefer one drive and have the csv file there you will have to change this line to connect to your one drive.
```{r inport the file}
#Daily <- terra::.csv(file.choose()) #option 1: if you dont know the file path address
Daily <- read.csv("Daily_temp.csv") #option 2: if you know the file path address
```

# Manpulating the data frame
Making the dates "true dates"  - this allows R to understand what data type this is and will read them as true dates. 
```{r 0}
#making the date column a true date
Daily <- mutate(Daily, Date = as.Date(Date, format= "%m/%d/%Y"))
```

Making a new column that is a month column based on the dates. If you ever try this code for a different data set you will have to adjust and edit the bounds of the dates and months included. The new Month column is a factor data type with 4 levels that are each month created.
```{r}
Daily$Month <- cut(Daily$Date, breaks = as.Date(c("2024-05-01", "2024-06-01", "2024-07-01", "2024-08-01", "2024-09-01")),
                    labels = c("May", "June", "July", "August"), right = FALSE)
```


# Dot graph

This is a good begining graph but doesn't tell us as much as we would like and could be more clear.
```{r 1.1}
DG1 <- ggplot(Daily, aes(x=Date, y=Temp_mean, color = Site_name)) +
  geom_point() +
  xlab("Date") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Tempature") +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
print(DG1) 
```


```{r 1.2}
DG2 <- ggplot(Daily, aes(x=Date, y=Temp_mean)) +
 # geom_line() +
  geom_point() +
 # stat_summary(aes(group = 1), fun.y = "mean", geom = "line") +
  stat_smooth(method = glm) +
  xlab("Date") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Tempature") +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
print(DG2) 
```

```{r 1.3}
DG3 <- ggplot(Daily, aes(x=Date, y=Temp_mean, color=Site_name)) +
 # geom_line() +
  geom_point() +
 # stat_summary(aes(group = 1), fun.y = "mean", geom = "line") +
  stat_smooth(method = glm, se=FALSE) +
  xlab("Date") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season 2 Tempature") +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
print(DG3) 
```

**1.3:** Temperature is divided by site in a grid, line of dots per recorded visit temp by date.

```{r 1.4}
DG4 <- ggplot(Daily, aes(x=Date, y=Temp_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "glm") +
    #method = "loess") +
  scale_y_continuous(limits = c(50, 90)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #stat_summary(aes(group = 1), fun.y = "mean", geom = "line") +
  xlab("Date") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Temperature") +
  scale_x_date(date_labels = "%b %d", date_breaks = "3 week") +
  facet_wrap(vars(Site_name))
print(DG4) 
```

Average documented tempature across all sites
```{r calc avg temp}
Avg_Temp <- Daily %>% 
  group_by(Date, Month) %>%
  summarise(Avgtemp = mean(Temp_mean, na.rm = TRUE))
```



```{r}
DG5 <- ggplot(Avg_Temp, aes(x=Date, y=Avgtemp)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") + #makes a more fitted curvey line
  #stat_summary(aes(group = 1), fun.y = "mean", geom = "line") +
  #stat_smooth(method = lm) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Field Season 2 Date (2024)") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Tempature") +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") 
print(DG5) 
```

# Box plots
**1.4:** Temperature is divided by all the months where all the sites are together. I think this is the **best** one.

```{r 1.5}
BP1 <- ggplot(Daily, aes(x=Month, y=Temp_mean)) +
  geom_boxplot() +
  xlab("Month") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Tempature")
print(BP1) 
```

**1.5:** Temperature is divided by month in each box plot and each site is seperate in the grid.

```{r 1.6}
BP2 <- ggplot(Daily, aes(x=Month, y=Temp_mean)) +
  geom_boxplot() +
  xlab("Month") +
  ylab("Temperature (Fahrenheit)") +
  ggtitle("2024 Field Season Tempature") +
  facet_wrap(vars(Site_name))
print(BP2) 
```
